\section{Signal Processing Methods}
\label{sec:sp}

This section provides the primary systematic contribution of this review: a taxonomy of signal processing methods applied to LCM of ball bearings, organised by processing family rather than by sensing modality. For each family the method is first described on its own terms -- its mathematical basis, assumptions, and outputs -- and then its specific applications across sensing modalities are discussed, drawing on the literature surveyed. A master cross-reference matrix linking SP method families to sensing modalities is provided in Table~\ref{tab:sp_crossref} at the end of this section.

The organisation of this section reflects the signal processing chain from raw sensor data to a lubrication condition estimate. Time-domain and frequency-domain methods (Sections~\ref{subsec:sp_time}--\ref{subsec:sp_freq}) operate directly on the measured signal. Time-frequency and decomposition methods (Section~\ref{subsec:sp_tf}) address non-stationarity. Signal enhancement methods (Section~\ref{subsec:sp_enhance}) improve the signal before feature extraction. Physics-based and model-driven methods (Section~\ref{subsec:sp_model}) incorporate lubrication physics to convert processed signals into physically interpretable quantities. Finally, machine learning and data fusion methods (Sections~\ref{subsec:sp_ml}--\ref{subsec:sp_fusion}) learn mappings from processed features to lubrication condition labels or regression targets.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Time-Domain Statistical Methods}
\label{subsec:sp_time}

Time-domain statistical methods extract scalar features from a windowed segment of the raw or bandpass-filtered sensor signal. They are computationally lightweight, robust to non-stationarity within the window, and directly interpretable in physical terms. These properties make them the most practically deployed class of methods in both research and industrial LCM applications.

\subsubsection{Standard Statistical Moments and Energy Metrics}
\label{subsubsec:moments}
The most widely reported features are:
\begin{itemize}
    \item \textbf{Root mean square (RMS):} $x_\mathrm{RMS} = \sqrt{\frac{1}{N}\sum_{i=1}^{N} x_i^2}$. Reflects the total signal energy over the window. For passive methods, RMS increases monotonically with worsening lubrication in the mixed and boundary regimes due to increasing asperity contact energy. Widely reported for AE \cite{hou_-situ_2025,cornel_condition_2021,yoshioka_monitoring_2009,zhang_health_2025,renhart_monitoring_2021}, vibration \cite{jakobsen_vibration_2021,jakobsen_detecting_2023,xu_vibration-based_2024}, and thermal monitoring \cite{moussa_passive_2017}.
    \item \textbf{Kurtosis:} $K = \frac{\frac{1}{N}\sum_{i=1}^{N}(x_i - \bar{x})^4}{\left(\frac{1}{N}\sum_{i=1}^{N}(x_i - \bar{x})^2\right)^2}$. Measures the impulsiveness of the signal relative to a Gaussian distribution (for which $K=3$). High kurtosis indicates sparse, high-amplitude transients such as those produced by asperity micro-contacts or incipient crack impacts. Usgame Sandoval et al.\ \cite{usgame_sandoval_acoustic_2013} report an $\sim$8.6-fold increase in AE kurtosis from normal to severe fault in tapered roller bearings. Yoshioka and Shimizu \cite{yoshioka_monitoring_2009} note that under grease lubrication, kurtosis of the raw vibration signal decreases as damage transitions from sparse impulsive AE to a continuous broadband noise floor -- a counter-intuitive reversal that illustrates the importance of bearing-specific calibration. Kurtosis is also used as a band-selection criterion in spectral kurtosis (Section~\ref{subsubsec:spectral_kurtosis}).
    \item \textbf{Crest factor:} peak-to-RMS ratio. Sensitive to impulsive events but tends to decrease as the signal transitions from sparse impulses to continuous broadband noise, similarly to kurtosis.
    \item \textbf{AE amplitude and energy centres (Hou \& Zhang):} Rather than a single scalar, Hou and Zhang \cite{hou_-situ_2025} define the amplitude centre $A_c = \frac{1}{N}\sum A_i$ and energy centre $E_c = \frac{1}{N}\sum E_i$ over a 10-minute window, and form the two-dimensional pair $(A_c, E_c)$. The joint distribution in this space exhibits spatially distinct clusters for boundary, mixed, and EHL lubrication, enabling regime classification without a hard threshold. A speed-dependent weighted power-law regression of the form
    \begin{equation}
        \Lambda_E = W_A \cdot f_A(A_c, n) + W_E \cdot f_E(E_c, n)
    \end{equation}
    combines the two features -- with $W_A, W_E$ linear in shaft speed $n$ -- to estimate the film thickness ratio $\Lambda$ with 89.3\% accuracy. This is the most physically grounded time-domain AE approach reported in the reviewed literature.
    \item \textbf{Pulse count rate (PCR):} the number of AE threshold crossings per unit time. Miettinen et al.\ \cite{miettinen_analysis_2001} show that PCR increases monotonically with grease starvation and distinguishes ten grease formulations by thickener type and base-oil viscosity. PCR is the computational minimal AE processing approach and forms the basis of the industrial threshold methods (Section~\ref{subsubsec:threshold}).
    \item \textbf{Ring-down count:} the total number of threshold crossings within a single AE burst event. Usgame Sandoval et al.\ \cite{usgame_sandoval_acoustic_2013} report a $\sim$266-fold increase in ring-down count from normal to severe fault -- the most sensitive single time-domain feature in their study, outperforming kurtosis and RMS.
\end{itemize}

\textbf{Speed and load normalisation.} Time-domain features are sensitive to changes in operating speed and load independently of lubrication condition. This confounding effect is addressed by normalising features against a speed-dependent or load-dependent baseline: Cornel et al.\ \cite{cornel_condition_2021} normalise AE parameters against reference measurements at the same operating point; Hou and Zhang \cite{hou_-situ_2025} explicitly incorporate speed as a predictor in their film-thickness regression. Without such normalisation, a feature increase caused by a speed change is indistinguishable from a lubrication-related change, which limits the applicability of fixed-threshold industrial methods to constant-speed machinery.

\subsubsection{Threshold Methods}
\label{subsubsec:threshold}
The simplest operational approach is to compare a time-domain feature -- most commonly the ultrasound dB level or AE RMS -- against a threshold established during a reference measurement under known adequate lubrication. If the feature rises a specified number of dB above the reference, relubrication is recommended \cite{lineham_ultrasonic_2008}. This is the method employed by industrial instruments such as UE Systems Ultraprobe and Sonotec's handheld devices. The advantages are simplicity, interpretability, and no requirement for labelled training data. The critical disadvantage is that the threshold carries no information about operating condition: a speed or load change will trigger a false alarm, and gradual lubrication degradation that produces a slow trend may not exceed the threshold until damage is well developed.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Frequency-Domain Methods}
\label{subsec:sp_freq}

Frequency-domain methods transform the time-domain signal into a spectral representation, enabling identification of narrowband features associated with periodic bearing kinematics superimposed on broadband tribological noise.

\subsubsection{Power Spectral Density and Bearing Characteristic Frequencies}
\label{subsubsec:psd}
The Fourier transform and power spectral density (PSD) reveal the distribution of signal energy across frequency. In the context of LCM, two types of spectral content are of interest: (1) broadband increases in spectral energy associated with increased asperity contact under inadequate lubrication, and (2) narrowband peaks at bearing characteristic frequencies -- the ball pass frequency of the outer race (BPFO), ball pass frequency of the inner race (BPFI), ball spin frequency (BSF), and fundamental train frequency (FTF) -- which indicate rolling element excitation of resonances in the bearing structure. Jakobsen et al.\ \cite{jakobsen_vibration_2021} use PSD features of vibration and ultrasound signals as inputs to regression models predicting $\kappa$, demonstrating that frequency-domain features carry lubrication-related information beyond what is captured by time-domain statistics alone.

\subsubsection{Envelope Analysis}
\label{subsubsec:envelope}
Envelope analysis demodulates a high-frequency carrier signal -- such as a structural resonance excited by bearing impacts -- to extract the low-frequency amplitude modulation produced by periodic contact events \cite{randall_vibrationbased_2021}. The approach consists of: (1) bandpass filtering around a resonance frequency to isolate the carrier, (2) computing the analytic signal via the Hilbert transform to obtain the instantaneous envelope, and (3) computing the FFT of the envelope (the envelope spectrum) to reveal modulation frequencies. In the LCM context, envelope analysis of AE signals at $>$100\,kHz reveals BPFO and its harmonics in the envelope spectrum before they appear in conventional vibration measurements. Yoshioka and Shimizu \cite{yoshioka_monitoring_2009} apply this approach to AE from a deep-groove ball bearing and detect BPFO peaks approximately 40\,000 revolutions before vibration RMS increases. The same principle is used by Kestel et al.\ \cite{kestel_condition_2025} for AE envelope analysis in multi-sensor LCM.

\subsubsection{Spectral Kurtosis and Optimal Band Selection}
\label{subsubsec:spectral_kurtosis}
Spectral kurtosis (SK) quantifies the kurtosis of the signal's short-time Fourier transform (STFT) coefficient distribution at each frequency bin, identifying frequency bands where impulsive transients have the highest signal-to-noise ratio. The frequency band with the highest SK can be used to design an optimal bandpass filter for envelope analysis. Jakobsen et al.\ \cite{jakobsen_detecting_2023} apply SK to vibration and ultrasound signals and find that the SK-optimal band is correlated with $\kappa$ values, providing a principled frequency-domain LCM indicator that adapts to the specific bearing and operating condition.

\subsubsection{Fractional-Octave Filter Banks}
\label{subsubsec:octave}
Rather than a single broadband metric, fractional-octave filter banks decompose the signal into a set of frequency bands following a geometric frequency progression. Renhart et al.\ \cite{renhart_monitoring_2021} apply a 1/12-octave filter bank (spanning approximately 50--425\,kHz) to AE signals from journal bearing segments. Within each band and time window, they compute the composite parameter $P_5 = x_\mathrm{RMS} \cdot K$ -- the product of RMS and kurtosis -- which is sensitive to transient impulsive events that are simultaneously energetic and non-Gaussian. The approach reveals mechanistic discrimination unavailable from broadband time-domain statistics: the 86.8\,kHz band is dominant during adhesive wear at low speeds, while the 206--232\,kHz bands increase at higher speeds and are attributed to particle-driven wear, following the tribological wear mechanism classification of Hase et al.\ This frequency-resolved mechanistic discrimination is a key advantage of spectral processing for LCM.

\subsubsection{Multi-Band Hit-Rate and Wear Particle Index}
\label{subsubsec:hitrate}
The Delft University group \cite{scheeren_acoustic_2024,delft_university_of_technology_tu_delft_influence_2024} employ a three-band sensor array (40--100\,kHz, 95--180\,kHz, 180--580\,kHz) for AE monitoring of large-scale low-speed slew bearings. An AE hit is recorded when the signal in a given band exceeds a threshold with a minimum SNR of 20\,dB, and the hit rate is normalised by loading cycles to remove the effect of varying test duration. A Wear Particle Index (WPI) derived from the normalised hit rate shows linear correlation with abrasive particle mass fraction in grease. Crucially, non-abrasive (soft) particles produce negligible additional hits, demonstrating that AE hit-rate is selective for hardness contrast rather than total particle mass -- a practically important limitation for contamination monitoring.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Time-Frequency and Decomposition Methods}
\label{subsec:sp_tf}

Time-frequency methods represent signal energy jointly in time and frequency, addressing the non-stationarity that characterises LCM signals arising from variable operating conditions or transient lubrication events.

\subsubsection{Short-Time Fourier Transform}
\label{subsubsec:stft}
The STFT divides the signal into short, overlapping windows and computes the FFT of each, yielding a spectrogram. The time-frequency resolution is governed by the window length (longer windows yield better frequency resolution but poorer time resolution). For LCM, the spectrogram is used to track the evolution of spectral features over time and to visualise non-stationary changes associated with lubrication degradation. Grease contamination detection via vibration spectrogram features combined with CNN-VGG16 classification is demonstrated in \cite{kannan_review_2024}. The STFT is also used within the SK framework as the basis for computing the kurtogram \cite{jakobsen_detecting_2023,kestel_condition_2025}.

\subsubsection{Wavelet Transform}
\label{subsubsec:wavelet}
The continuous wavelet transform (CWT) provides time-frequency resolution that adapts to the local signal frequency through a scaling parameter: short wavelets at high frequencies and long wavelets at low frequencies, giving better time resolution where signals are fast-varying and better frequency resolution where they are slow-varying. This property is well-suited to AE bursts, which are transient and broadband. Renhart et al.\ \cite{renhart_monitoring_2021} apply the CWT with a Morse (generalised Morse) wavelet to individual AE bursts triggered by the $P_5$ criterion, producing amplitude scalograms that reveal a two-stage within-event mechanism: an initial high-frequency asperity impact followed by a lower-frequency relaxation. Adaptive wavelet denoising -- in which the wavelet decomposition threshold is optimised by a kurtosis-entropy criterion -- is applied to non-contact acoustic (microphone) signals from bearings to denoise and classify bearing health \cite{pacheco-cherrez_bearing_2022}.

The discrete wavelet transform (DWT) and its multi-resolution analysis framework are also applied to vibration signals for LCM, enabling simultaneous extraction of features at multiple frequency scales corresponding to different physical phenomena \cite{wang_advanced_2024,jakobsen_vibration_2021}.

\subsubsection{Empirical Mode Decomposition and Hilbert--Huang Transform}
\label{subsubsec:emd}
Empirical mode decomposition (EMD) adaptively decomposes a signal into intrinsic mode functions (IMFs) by iterative sifting of local extrema, without requiring a predefined basis. The Hilbert transform applied to each IMF yields instantaneous frequency and amplitude, producing the Hilbert spectrum -- the Hilbert--Huang Transform (HHT). EMD is particularly suited to non-linear and non-stationary signals, properties that characterise bearing signals under time-varying lubrication conditions \cite{wang_advanced_2024,jakobsen_vibration_2021}. However, EMD suffers from mode mixing in the presence of intermittent signals, which has motivated variants such as ensemble EMD (EEMD) and variational mode decomposition (VMD) \cite{jaros_advanced_2023}.

\subsubsection{Symplectic Geometry Mode Decomposition}
\label{subsubsec:sgmd}
Symplectic geometry mode decomposition (SGMD) is a signal decomposition method based on the symplectic geometry of the phase space of the signal's trajectory matrix. Yu et al.\ \cite{yu_novel_2021} apply SGMD to signals from inductive debris sensors, extracting debris-particle signatures from inductive bearing monitoring signals buried in noise and harmonic interference. Pan et al.\ \cite{pan_symplectic_2019} provide the methodological foundation. The method offers improved separation of overlapping frequency components compared to EMD-based methods in specific application scenarios.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Signal Enhancement Methods}
\label{subsec:sp_enhance}

Signal enhancement methods improve the signal-to-noise ratio or impulsiveness of the raw sensor data prior to feature extraction, without reducing the signal to a scalar feature.

\subsubsection{Minimum Entropy Deconvolution}
\label{subsubsec:med}
Minimum entropy deconvolution (MED) designs a finite impulse response (FIR) filter that maximises the kurtosis (equivalently, minimises the entropy) of the filter output, thereby recovering the sparse impulsive excitation source from a convolved, noisy observation. In the LCM context, the impulsive sources are the transient mechanical impacts associated with asperity contacts or rolling element passage over roughened raceways under starved lubrication. Xu et al.\ \cite{xu_vibration-based_2024} demonstrate that MED applied to vibration signals significantly enhances the detection of lubrication starvation-related impulses, enabling spectral centroid and envelope features to detect starved conditions that are not detectable in the unprocessed signal. The method assumes that the source signal is sparse (impulsive), which may not hold for all lubrication fault types. Multipoint optimal minimum entropy deconvolution adjusted (MOMEDA) and its variants extend MED to multi-channel and periodic-source scenarios \cite{jaros_advanced_2023}.

\subsubsection{Resonance-Based Sparse Signal Decomposition}
\label{subsubsec:rssd}
Resonance-based sparse signal decomposition (RSSD) decomposes the signal into a high-resonance (sustained, oscillatory) component and a low-resonance (impulsive) component using morphological component analysis with Q-factor wavelet transform bases. The impulsive component captures the bearing contact events relevant to LCM. Herwig et al.\ \cite{herwig_signal_2025} demonstrate RSSD combined with recursive least squares (RLS) de-reverberation for non-contact acoustic (microphone) bearing monitoring, reducing the masking effect of multi-path acoustic reflections on the impulsive bearing signal. The approach is relevant to passive, non-contact sensing where channel effects and reverberation are significant.

\subsubsection{Waveform-Similarity Clustering}
\label{subsubsec:clustering}
For low-speed bearings ($<$20\,rpm) where individual AE events are temporally resolvable, Scheeren et al.\ \cite{scheeren_acoustic_2024} develop a waveform-similarity clustering method based on pairwise cross-correlation. The cross-correlation coefficient between two signals $s_i$ and $s_j$ is used to group AE hits into clusters of consistent waveform shape, exploiting the fact that a fixed AE source (a localised raceway defect) produces consistently similar waveforms at a given sensor. An additional multi-channel coherence criterion rejects sources that do not produce consistent time delays across the sensor array. Clusters with intra-cluster similarity $>$93\% and hit-rates of 1--2 per loading cycle were identified at $\sim$170\,000 loading cycles and confirmed by post-test surface inspection. The method is the most powerful approach reviewed for isolating a localised defect against a continuous background of asperity-contact emissions at low speeds, but it requires dense sensor arrays, offline computation, and processes $\sim$2.3 million signals per experiment -- currently unsuitable for real-time deployment.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Physics-Based and Model-Driven Methods}
\label{subsec:sp_model}

Physics-based methods incorporate lubricant or structural mechanics models to convert processed signals into physically interpretable quantities such as film thickness or lubrication regime.

\subsubsection{Quasi-Static Spring-Layer Model for Ultrasonic Reflectometry}
\label{subsubsec:spring}
Active ultrasonic reflectometry measures the reflection coefficient $R$ of an ultrasonic pulse at the raceway--lubricant interface. For a thin lubricant film (film thickness much less than the acoustic wavelength), the spring-layer model relates $R$ to the film stiffness $K^*$:
\begin{equation}
    R = \frac{1}{1 + \frac{2Z_1 K^*}{\omega (Z_1^2 - Z_2^2)}} \cdot \left(\frac{Z_2 - Z_1}{Z_2 + Z_1}\right),
\end{equation}
where $Z_1$, $Z_2$ are the acoustic impedances of the raceway and lubricant materials, $\omega$ is the angular frequency of the ultrasonic pulse, and $K^*$ is the complex stiffness of the lubricant layer related to film thickness $h$ and lubricant bulk modulus $E^*$ by $K^* = E^*/h$ \cite{drinkwater_ultrasonic_2009,dou_review_2022}. This model enables direct, calibrated film thickness estimation from the measured reflection coefficient. Validation against Hamrock--Dowson predictions is reported in \cite{dwyer-joyce_measurement_2003,zhang_ultrasonic_2015,chen_evaluation_2023}. Phase-based algorithms that improve accuracy for thin films are described in \cite{wang_accuracy-improved_2024}. Sub-Nyquist sampling to reduce data acquisition requirements while retaining film thickness accuracy is demonstrated by Chen et al.\ \cite{chen_sub-nyquist_2025}. Li et al.\ \cite{li_lubrication_2026} extend the spring-layer model to the curved contact geometry of wind turbine pitch bearings using ray-modelling corrections for transducer curvature and surface geometry.

\subsubsection{Electrical Circuit Models}
\label{subsubsec:electrical_model}
For electrical sensing methods (capacitance, impedance, resistance), physics-based circuit models relate the measured electrical signal to the lubricant film thickness. The bearing is modelled as a parallel combination of capacitive and resistive elements whose values depend on the film thickness, lubricant dielectric constant, and Hertzian contact geometry. Cen et al.\ \cite{cen_ehl_2021} and Becker-Dombrowsky et al.\ \cite{becker-dombrowsky_electrical_2024} review the capacitance and impedance formulations respectively. Maruyama et al.\ \cite{maruyama_lubrication_2023} apply the impedance model to practical deep-groove ball bearings under operating conditions. Two approaches to model calibration are reported: purely physics-based models using known material properties, and empirical calibration tests using reference lubricants to derive a relationship between the measured signal and Hamrock--Dowson-predicted film thickness \cite{cen_ehl_2021,schneider_method_2022}.

\subsubsection{Kappa and Lambda as Regression Targets}
\label{subsubsec:kappa_target}
A model-driven approach for passive sensing methods uses $\kappa$ or $\Lambda$ -- computed from the Hamrock--Dowson model and operating parameters -- as the regression target for data-driven models trained on sensor features. This provides a physically grounded target variable linked to ISO standards without requiring invasive lubricant sampling. Jakobsen et al.\ \cite{jakobsen_detecting_2023} train LASSO regression and neural network models on time-domain and spectral-kurtosis features from vibration and passive ultrasound signals to predict $\kappa$, and demonstrate correlation between predicted $\kappa$ and controlled lubrication conditions across multiple bearing configurations. The approach effectively bridges the gap between physics-based lubrication models and data-driven signal processing.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Machine Learning and Deep Learning}
\label{subsec:sp_ml}

Data-driven methods learn mappings from handcrafted or raw signal features to lubrication condition labels or continuous condition estimates. They can be organised into classical machine learning (operating on handcrafted features), deep learning (operating on raw signals or spectrograms), and emerging self-supervised approaches.

\subsubsection{Classical Machine Learning}
\label{subsubsec:classical_ml}
Classical machine learning methods -- support vector machines (SVM), random forests, logistic regression, and regularised linear regression (LASSO) -- are applied to vectors of handcrafted features extracted by the time-domain, frequency-domain, or time-frequency methods described above. Jakobsen et al.\ \cite{jakobsen_detecting_2023} demonstrate that LASSO regression on vibration and ultrasound features achieves good $\kappa$ prediction while providing interpretable feature weights. Wang et al.\ \cite{wang_evaluation_2018} apply grey-correlation K-means clustering to oil multi-parameter monitoring data (viscosity, density, permittivity, water content, debris) to evaluate bearing wear states over full bearing lifetime, achieving earlier failure prediction than vibration signals alone. Classification of lubrication fault types (starvation vs.\ contamination vs.\ degraded viscosity) using SVM with handcrafted AE or vibration features is demonstrated in \cite{kestel_condition_2025,cornel_condition_2021}.

\subsubsection{Deep Learning}
\label{subsubsec:dl}
Deep learning methods operate on raw signals or intermediate representations (spectrograms, graph adjacency matrices) and learn features jointly with the classification or regression task. Key architectures applied to LCM include:
\begin{itemize}
    \item \textbf{Convolutional neural networks (CNNs):} applied to lubricant debris images for particle classification \cite{wang_online_2022}, and to vibration spectrograms for grease contamination detection \cite{kannan_review_2024}. Zhang et al.\ \cite{zhang_health_2025} apply a vision-language large language model (PaLiGemma) to images of AE time-domain waveforms for automated wear detection, reporting higher classification accuracy than conventional threshold methods.
    \item \textbf{Graph neural networks (GNNs):} Zhang et al.\ \cite{zhang_graph-data-based_2024} represent multi-sensor data (vibration, AE, temperature) as a graph and apply graph convolutional operations to detect lubrication degradation, leveraging sensor correlations that are not captured by independent feature extraction.
    \item \textbf{Recurrent and hybrid architectures:} temporal dependencies in lubrication degradation signals can be captured by LSTM and CNN-LSTM hybrid models, relevant to monitoring slowly evolving degradation trends \cite{hotait_intelligent_2021}.
\end{itemize}
A consistent challenge for deep learning in LCM is the scarcity of labelled training data from real bearing degradation under controlled lubrication conditions. Controlled accelerated tests are expensive to run, and publicly available LCM datasets with varied lubrication conditions and sensing modalities are not currently available.

\subsubsection{Self-Supervised and Foundation Model Approaches}
\label{subsubsec:ssl}
Self-supervised learning (SSL) methods pre-train representations on large amounts of unlabelled time series data by solving auxiliary tasks (contrastive learning, masked autoencoding, time-frequency consistency) before fine-tuning on small labelled datasets \cite{shamba_contrast_2024}. The time-frequency consistency framework (TF-C) \cite{siddique_hybrid_2025} demonstrates transfer of pre-trained representations to bearing fault detection tasks with limited labels, which is directly relevant to LCM where labelled degradation data is scarce. Foundation model approaches -- large pre-trained models applied to time series with zero-shot or few-shot adaptation \cite{sun_state---art_2025} -- represent an emerging direction for bearing monitoring, though direct application to LCM has not yet been demonstrated in the reviewed literature. These approaches are discussed further as a research gap in Section~\ref{sec:gaps}.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Multi-Sensor Data Fusion}
\label{subsec:sp_fusion}

Multi-sensor fusion combines information from multiple sensing modalities to obtain lubrication condition estimates more accurate or robust than any single modality alone.

\subsubsection{Feature-Level Fusion}
Feature-level (early) fusion concatenates feature vectors from multiple sensors into a single feature vector before the classification or regression step. Zhang et al.\ \cite{zhang_graph-data-based_2024} demonstrate feature fusion of vibration, AE, and temperature signals for lubrication degradation detection, showing that the fused system outperforms any single-modality approach. Zhao et al.\ \cite{zhao_fault_2023} combine oil debris features with vibration features for bearing fault diagnosis, reporting a 15.7\% improvement in diagnostic accuracy from the oil debris features alone. Cornel et al.\ \cite{cornel_condition_2021} combine AE with vibration for complementary early detection under varying lubrication conditions.

\subsubsection{Multi-Target Tracking for Debris}
For optical debris monitoring, multi-target tracking algorithms combine blob extraction from debris images with Kalman filtering to track individual debris particles and compute their velocity, which is correlated with debris-related bearing damage \cite{liu_oil_2022}. The Kalman filter provides temporal smoothing and handles particle occlusion.

\subsubsection{Multi-Parameter Oil Monitoring}
Wang et al.\ \cite{wang_evaluation_2018} combine six independently measured oil parameters (viscosity, density, permittivity, water content, temperature, ferrous debris count) using a grey-correlation model, aggregating diverse physical information into a scalar wear state indicator. This multi-parameter approach is analogous in structure to feature-level fusion of passive sensor signals, and demonstrates that the information from multiple oil quality indicators is complementary for classifying bearing wear stages.

% ─────────────────────────────────────────────────────────────────────────────
\subsection{Master Cross-Reference Matrix}
\label{subsec:sp_crossref}

Table~\ref{tab:sp_crossref} summarises the application of SP method families to each sensing modality, as found in the reviewed literature. A filled cell indicates the combination has been demonstrated for LCM in at least one reviewed study.

\begin{table}[H]
\scriptsize
\centering
\renewcommand{\arraystretch}{1.4}
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\linewidth}{|
>{\raggedright\arraybackslash}p{3cm}
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X
| >{\centering\arraybackslash}X |}
\hline
\textbf{SP Method Family}
& \rotatebox{75}{\textbf{Vibration}}
& \rotatebox{75}{\textbf{Acoustic Emission}}
& \rotatebox{75}{\textbf{Ultrasound (passive)}}
& \rotatebox{75}{\textbf{Sound / Microphone}}
& \rotatebox{75}{\textbf{Thermal}}
& \rotatebox{75}{\textbf{Active Ultrasonic}}
& \rotatebox{75}{\textbf{Electrical}}
& \rotatebox{75}{\textbf{Magnetic / Oil}}
\\
\hline
Time-domain statistics      & \checkmark & \checkmark & \checkmark &            & \checkmark &            &            & \checkmark \\
\hline
Threshold methods           &            & \checkmark & \checkmark &            &            &            &            &            \\
\hline
PSD / Bearing char.\ freqs  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            &            \\
\hline
Envelope analysis           & \checkmark & \checkmark &            &            &            &            &            &            \\
\hline
Spectral kurtosis           & \checkmark & \checkmark & \checkmark &            &            &            &            &            \\
\hline
Octave filter bank          &            & \checkmark &            &            &            &            &            &            \\
\hline
STFT / Spectrogram          & \checkmark & \checkmark &            &            &            &            &            &            \\
\hline
Wavelet transform           & \checkmark & \checkmark &            & \checkmark &            &            &            &            \\
\hline
EMD / HHT                   & \checkmark & \checkmark &            &            &            &            &            &            \\
\hline
Mode decomposition (SGMD)   &            &            &            &            &            &            &            & \checkmark \\
\hline
MED / Deconvolution         & \checkmark & \checkmark &            &            &            &            &            &            \\
\hline
RSSD / Sparse decomposition &            &            &            & \checkmark &            &            &            &            \\
\hline
Waveform clustering         &            & \checkmark &            &            &            &            &            &            \\
\hline
Spring-layer model          &            &            &            &            &            & \checkmark &            &            \\
\hline
Electrical circuit model    &            &            &            &            &            &            & \checkmark &            \\
\hline
$\kappa$/$\Lambda$ regression target & \checkmark & \checkmark & \checkmark &  &            &            &            &            \\
\hline
Classical ML (SVM, LASSO)   & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \checkmark \\
\hline
Deep learning (CNN, GNN)    & \checkmark & \checkmark &            & \checkmark & \checkmark &            &            &            \\
\hline
Self-supervised / SSL       & \checkmark &            &            &            &            &            &            &            \\
\hline
Multi-sensor fusion         & \checkmark & \checkmark &            &            & \checkmark &            &            & \checkmark \\
\hline
\end{tabularx}
\caption{Master cross-reference matrix of signal processing method families and sensing modalities for LCM of ball bearings. A \checkmark indicates the combination has been demonstrated for LCM in at least one reviewed study. Empty cells represent combinations that are theoretically applicable but not yet demonstrated in the reviewed literature.}
\label{tab:sp_crossref}
\end{table}
